---
title: "Youth Mental Health & Social Media"
author: "Dean Marchiori"
format: html
---

# Mental Health Conditions by Age & Gender  

Setup and load packages. 

```{r}
library(tidyverse)
library(themes360info)  #360Info internal package: `remotes::install_github("360-info/themes360info")` 
```

## Import Data  

Download and unzip data. For more information see `data/README.md`.  

```{r}
# download.file("https://www.abs.gov.au/census/find-census-data/datapacks/download/2021_GCP_AUS_for_AUS_short-header.zip",
#     destfile = "data/ABS_Aus_2021.zip"
# )

# unzip("data/ABS_Aus_2021.zip", exdir = "data/ABS_Aus_2021")
```

Read in the relevant table from the ABS GCP datasets. Table 19 in the GCP is around long term health conditions. 
It is split across three csv files (A, B, C).  

```{r}
mh19a <- read_csv("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19A_AUS_AUS.csv")
mh19b <- read_csv("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19B_AUS_AUS.csv")
mh19c <- read_csv("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19C_AUS_AUS.csv")
```

## Data Processing  

Extracting Male and Female Mental Health data, and population totals.  

Pivoting wider and separating out the ABS column names to decompose gender and age .  

I have calculated *_directional columns to allow plotting like a symmetric age pyramid.  

```{r}

# This is a helper function for the symmetric bar plot
pretty_symmetric <- function(range, n = 5) {
    range_1 <- c(-range[1], range[2])
    range_2 <- c(range[1], -range[2])

    pretty_vec_1 <- pretty(range_1)
    pretty_vec_2 <- pretty(range_2)

    pretty(
        c(pretty_vec_1, pretty_vec_2),
        n = n
    )
}

# data processing pipeline
mh <- bind_cols(mh19a, mh19b, mh19c) |>
    select(
        AUS_CODE_2021 = AUS_CODE_2021...1,
        contains(c("M_Mental", "F_Mental", "M_Tot", "F_Tot")),
        -M_Mental_health_cond_Tot,
        -F_Mental_health_cond_Tot,
        -F_Tot_Tot,
        -M_Tot_Tot
    ) |>
    pivot_longer(-AUS_CODE_2021,
        names_to = "cat"
    ) |>
    separate_wider_regex(cat,
        patterns = c(gender = ".", cond = "_.+", age = "(?<=_)[0-9]+_.+$"),
        too_few = "align_start"
    ) |>
    pivot_wider(names_from = cond) |>
    rename(
        condition = `_Mental_health_cond_`,
        total = `_Tot_`
    ) |>
    mutate(
        age = str_replace(age, "_", "-"),
        gender = fct_recode(gender, Female = "F", Male = "M"),
        alpha_channel = ifelse(age %in% c("15-24", "0-14", "25-34"), 0.8, 0.6)
    ) |>
    mutate(
        age = fct_rev(age),
        condition_directional = ifelse(gender == "Male", condition * -1, condition),
        proportion = condition / total,
        proportion_directional = ifelse(gender == "Male", proportion * -1, proportion)
    )

```


## Plot  

Showing symmetric age pyramid, split by gender for Mental Health prevalance. 

This version shows **count of persons**  

```{r}
abs_plot_1 <- ggplot(mh, aes(condition_directional, age, fill = gender)) +
    geom_col(aes(alpha = alpha_channel)) +
    scale_x_continuous(
        breaks = pretty_symmetric(range(mh$condition_directional)),
        labels = scales::comma(abs(pretty_symmetric(range(mh$condition_directional))))
    ) +
    expand_limits(x = pretty_symmetric(range(mh$condition_directional))) +
    scale_fill_manual(values = c("#36a7e9", "#2d3494"))+
    scale_alpha(range = c(0.6, 1)) +
    guides(alpha = "none") +
    themes360info::theme_360() +
    labs(
        title = "Persons with a mental health condition",
        subtitle = "Has been told by a doctor or nurse that they have a **mental health condition**<br>*(including depression or anxiety)*",
        caption = "**Source**: Based on Australian Bureau of Statistics data.",
        x = "Count of Persons",
        y = "Age",
        fill = "Gender"
    ) +
    annotate_360_lightblue(
        x = 210000, y = 9.1, size = 3, 
        label = "**Young people** make up a<br> large proprtion of overall sufferers<br> of mental health conditions.<br> **Females** are much more heavily impacted."
    ) 

save_360plot(abs_plot_1, "out/abs_plot_1.png", shape = "square")

abs_plot_1
```

```{r}
ggplot(mh, aes(proportion_directional, age, fill = gender)) +
    geom_col() +
    scale_x_continuous(
        breaks = pretty_symmetric(range(mh$proportion_directional)),
        labels = abs(pretty_symmetric(range(mh$proportion_directional)))
    ) +
    expand_limits(x = pretty_symmetric(range(mh$proportion_directional))) 
```


# Mental Health Map  

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(sfdep)
```

```{r}
# unzip("data/Geopackage_2021_G19_AUST_GDA2020.zip", exdir = "data/spatial")
```

```{r}
mh2 <- sf::read_sf("data/spatial/G19B_AUST_GDA2020.gpkg", layer = "G19B_SA2_2021_AUST")
mh3 <- sf::read_sf("data/spatial/G19C_AUST_GDA2020.gpkg", layer = "G19C_SA2_2021_AUST")
mh_all <- cbind(mh2, mh3)
```


```{r}
mh_dat <- mh_all |>
    st_drop_geometry() |>
    as_tibble()  |>
    select(
        SA2_CODE_2021,
        contains(c("P_Mental", "P_Tot")),
        -P_Tot_Tot,
        -P_Mental_health_cond_Tot
    ) |>
    pivot_longer(-c(SA2_CODE_2021),
        names_to = "cat"
    ) |>
    separate_wider_regex(cat,
        patterns = c(cond = "P_.+", age = "(?<=_)[0-9]+_.+$"),
        too_few = "align_start"
    ) |>
    pivot_wider(names_from = cond) |>
    rename(
        condition = P_Mental_health_cond_,
        total = P_Tot_
    ) |>
    mutate(
        prop = condition / total,
        prop = replace_na(prop, 0),
        prop = ifelse(condition > total, 1, prop)
    )  |>
    filter(age == "15_24")


mh_map <- mh_all |>
    select(SA2_CODE_2021, SA2_NAME_2021, geom) |>
    inner_join(mh_dat)  |>
    as_tibble()  |>
    st_as_sf()

```

```{r}
# mapview(mh_map, zcol = "prop")
```


```{r}
has_nb <- mh_map |>
    filter(
        !st_is_empty(geom)
    ) |>
    mutate(nb = st_contiguity(geom)) |>
    slice(which(map_int(nb, \(x) sum(x)) != 0))


mh_wts <- has_nb |>
    mutate(
        nb = st_contiguity(geom),
        wt = st_weights(nb)
    )

mh_wts |>
    st_drop_geometry() |>
    mutate(lag = st_lag(prop, nb, wt)) |>
    ggplot(aes(prop, lag)) +
    geom_point() +
    geom_vline(aes(xintercept = median(prop)), lty = 2) +
    geom_hline(aes(yintercept = median(lag)), lty = 2) +
    geom_smooth(method = "lm")


mh_moran <- mh_wts |>
    mutate(
        local_moran = local_moran(prop, nb, wt, nsim = 199),
    ) |>
    tidyr::unnest(local_moran) |>
    mutate(mean = ifelse(p_folded_sim <= 0.05, as.character(mean), NA))




```

```{r}
out <- ggplot(mh_moran, aes(fill = mean)) +
    geom_sf() +
    theme_void() 
plotly::ggplotly(out)
```

```{r}


hh <- mh_moran  |>
filter(mean == "High-High")

mapviewOptions(fgb = FALSE)

    map <- leaflet(hh) |>
        leaflet::addProviderTiles(providers$CartoDB.Positron)

    mout <- mapview(hh, map = map)

    mapshot(mout, file = "test.png")
```


```{r}
p <- function (n) {
    grDevices::hcl.colors(n, palette = "viridis", rev = T)
}


mh_moran  |>
select(SA2_NAME_2021, condition, total, prop, mean)  |>
mapview(zcol = "mean", viewer.suppress = TRUE)


```
