---
title: Mental Health Conditions by Age & Gender 
subtitle: Analysis and data preparation of data from 2021 ABS Census. 
format:
  360-analysis-html: default
author: Dean Marchiori
date: last-modified
code-fold: true
---

# Mental Health Conditions by Age & Gender  

Setup and load packages. 

```{r}
library(tidyverse)
library(here)
library(themes360info)  #360Info internal package: `remotes::install_github("360-info/themes360info")` 
```

## Import Data  

Download and unzip data. For more information see `./data/README.md`.  

```{r}
if (!file.exists(here("data/ABS_Aus_2021.zip"))) {

    download.file("https://www.abs.gov.au/census/find-census-data/datapacks/download/2021_GCP_AUS_for_AUS_short-header.zip",
        destfile = here("data/ABS_Aus_2021.zip")
    )

    unzip(here("data/ABS_Aus_2021.zip"), exdir = here("data/ABS_Aus_2021"))
}
```

Read in the relevant table from the ABS GCP datasets. Table 19 in the GCP is around long term health conditions. 
It is split across three csv files (A, B, C).  

```{r}
mh19a <- read_csv(here("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19A_AUS_AUS.csv"))
mh19b <- read_csv(here("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19B_AUS_AUS.csv"))
mh19c <- read_csv(here("data/ABS_Aus_2021/2021 Census GCP Australia for AUS/2021Census_G19C_AUS_AUS.csv"))
```

## Data Processing  

Extracting Male and Female Mental Health data, and population totals.  

Pivoting wider and separating out the ABS column names to decompose gender and age.  

I have calculated *_directional columns to allow plotting like a symmetric age pyramid.  

```{r}
# data processing pipeline
mh <- bind_cols(mh19a, mh19b, mh19c) |>
    select(
        AUS_CODE_2021 = AUS_CODE_2021...1,
        contains(c("M_Mental", "F_Mental", "M_Tot", "F_Tot")),
        -M_Mental_health_cond_Tot,
        -F_Mental_health_cond_Tot,
        -F_Tot_Tot,
        -M_Tot_Tot
    ) |>
    pivot_longer(-AUS_CODE_2021,
        names_to = "cat"
    ) |>
    separate_wider_regex(cat,
        patterns = c(gender = ".", cond = "_.+", age = "(?<=_)[0-9]+_.+$"),
        too_few = "align_start"
    ) |>
    pivot_wider(names_from = cond) |>
    rename(
        condition = `_Mental_health_cond_`,
        total = `_Tot_`
    ) |>
    mutate(
        age = str_replace(age, "_", "-"),
        gender = fct_recode(gender, Female = "F", Male = "M"),
        alpha_channel = ifelse(age %in% c("15-24", "0-14", "25-34"), 0.8, 0.6)
    ) |>
    mutate(
        age = fct_rev(age),
        condition_directional = ifelse(gender == "Male", condition * -1, condition),
        proportion = condition / total,
        proportion_directional = ifelse(gender == "Male", proportion * -1, proportion)
    )

```

# Data Export 

Write out data as a csv  

```{r}
write_csv(mh, here("data/final/mental_health_census.csv"))
```