---
title: Abbreviated FOMO and social media dataset
subtitle: FoMO, but not self-compassion, moderates the link between social media use and anxiety in adolescence.   
format:
  360-analysis-html: default
author: Dean Marchiori
date: last-modified
code-fold: true
---

```{r}
library(tidyverse)
library(here)
library(themes360info)
```

From : Danielle A. Einstein, Carol Dabb & Madeleine Fraser (2023) FoMO, but not self-compassion, moderates the link between social media use and anxiety in adolescence, Australian Journal of Psychology, 75:1, DOI: 10.1080/00049530.2023.2217961  

> Social media use is ubiquitous during adolescence, and emerging research suggests an association with anxiety symptoms in some individuals. Two psychological constructs which may moderate this relationship are Fear of Missing Out (FoMO) and self-compassion. Higher FoMO tendencies may exacerbate the link between social media use and anxiety symptoms through greater fixation on social comparison, whereas higher self-compassion may weaken this link. The purpose of this study was to examine whether FoMO and self-compassion independently moderate the relationship between social media use and anxiety symptoms in adolescents.  

Read in the data from the study. For more info see `data/README.md`

```{r}
fomo <- read_csv(here("data/raw/Data file Einstein.csv"))
```

Adaptation of Figure 1 from the study.  

```{r}
lab <- str_wrap(c(
    "Once a week or less", "Several times a week", "About 1-4 times per day",
    "About 5-10 times per day", "More than 10 times a day", "Constantly"
), 10)

fomo_plot <- fomo |>
    select(SCAS_TOTAL_0, FoMO_Total, SM_0Reversed) |>
    mutate(
        fomo_hi_low = case_when(
            FoMO_Total <= 10.2 ~ "Average to Low FoMO",
            FoMO_Total >= 15.3 ~ "High FoMO"
        )
    ) |>
    drop_na() |>
    ggplot(aes(SM_0Reversed, SCAS_TOTAL_0, colour = fomo_hi_low)) +
    geom_point(position = position_jitterdodge(), aes(alpha = fomo_hi_low)) +
    geom_smooth(method = "lm", se = F) +
    scale_y_sqrt(
        breaks = c(0, 25, 50, 75, 114),
        labels = c(0, 25, 50, 75, 114),
        limits = c(0, 144)
    ) +
    scale_x_continuous(breaks = 1:6, labels = lab) +
    scale_color_manual(values = c("#26d4e0", "#cf1d1d")) +
    scale_alpha_manual(values = c(1, 0.9)) +
    themes360info::theme_360() +
    theme(
        legend.position = "none",
        axis.text.x = element_text(angle = 0, size = 10),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.subtitle = ggtext::element_markdown(lineheight = 1)
    ) +
    labs(
        title = "TEEN FOMO",
        subtitle = "A survey of 951 teenagers found that teens who reported a <span style='color:#cf1d1d'>**high fear of missing out**</span>, or FOMO,<br> also reported increased anxiety as they used social media more. But teens with <span style='color:#26d4e0'>**low FOMO**</span><br> reported less anxiety as they used it more.",
        caption = "**CHART:** James Goldie, 360info & Dean Marchiori, Wave Data Labs<br>**DATA:** Einstein et al. (2023) [doi: 10.1080/00049530.2023.2217961]",
        x = "Frequency of social media use",
        y = "Anxiety symptom score",
        colour = "Level of FoMO"
    )

fomo_plot

save_360plot(fomo_plot, here("out/img/fomo.png"), shape = "square")
```
